---
title: "Hypothesis testing"
output: 
  pdf_document:
    latex_engine: xelatex # or pdflatex if preferred
---

## Reading data and loading libraries
```{r warning=FALSE, message=FALSE}
# Load necessary libraries (if required)
# library(readr)
rm(list=ls())
library(ggplot2)
library(tidyverse)
# Read the CSV file 

data <- read.csv("anes_timeseries_2020_csv_20220210/anes_timeseries_2020_csv_20220210.csv")
```


## Data Wrangling 

For deciding if an observation belongs to a democrat or a republican, we use
V201226 PRE: SPLIT-TICKET VOTING, because...
lean = 1 $\rightarrow$ Republican
lean = 2 $\rightarrow$ Democrat

For the difficulty the decision variable is 
V202119 POST: HOW DIFFICULT WAS IT FOR R TO VOTE, since ...
difficulty = 1 $\rightarrow$ No difficult
difficulty $> 1 \rightarrow$ Difficult



```{r warning=FALSE, message=FALSE }
selected_data <- data %>%
  select(difficulty='V202119', lean='V201226')
filtered <- filter(selected_data, lean %in% c(1, 2), 
                   difficulty %in% c(1, 2, 3, 4, 5))
input_data <-filtered %>%
  mutate(difficulty = ifelse(difficulty == 1, 0, 1))
```

## PLOTS

```{r warning=FALSE, message=FALSE }
ggplot(input_data, aes(x = factor(difficulty), fill = factor(difficulty))) +
  geom_bar() +
  facet_wrap(~ factor(lean, labels = c("1 - Republican", "2 - Democrat")), ncol = 1) +
  labs(title = "Bar Plots by Party ",
       x = "Difficulty",
       y = "Count") +
  scale_x_discrete(labels = c("1" = "Difficult", "0" = "No difficult")) +
  scale_fill_discrete(name = "Difficulty",
                      labels = c("0" = "No difficult", "1" = "Difficult")) +
   theme_minimal()
```
 ## Hypothesis testing

$H_0:$ There is no difference between the Republicans and Democrats in terms
       of the difficulty when voting
       
```{r warning=FALSE, message=FALSE }
# Create a contingency table
contingency_table <- table(input_data$lean, input_data$difficulty)

# Perform a chi-squared test
chi_squared_test <- chisq.test(contingency_table)

# Print the results
print(chi_squared_test)
```
## Conclusions

Since p-value<0.05 there is enough statistical evidence to reject the null 
hypothesis, and we can conclude that there is difference on how Republicans and 
Democrats conceive difficulty when voting.
 
```{r warning=FALSE, message=FALSE }
total_counts <- rowSums(contingency_table)
percentage_difficulty <- (contingency_table[,2] / total_counts) * 100

# Add percentages as a new column to the contingency table
contingency_table_with_percentages <- cbind(contingency_table, Percentage_Difficulty = percentage_difficulty)

# Print the contingency table with percentages
contingency_table_with_percentages

```
The table above shows that in proportion the Republicans find more difficult to
vote, the sum of this result and the latter let us conclude that the Republicans
find more difficulty when voting.